NumPy---Array Computation for Python

### 2019-12-11 paper.tex
First shot at 'Library Organization'

\subsection*{Library organisation}

>>>>

\subsection*{Library organisation}
Broadly, the NumPy library consists of the following parts:
the NumPy array data structure; the so-called \emph{universal functions};
a set of library functions for manipulating arrays and doing scientific
computation; infrastructure libraries for unit tests and Python package
building; and the program \code{f2py} for wrapping Fortran code in Python.
The \code{ndarray} and the universal functions are generally considered
the core of the library.
In the following, we give a brief summary of these components of the
library.

\paragraph{\emph{Core.}}  The \code{ndarray} data structure and the
universal functions make up the core of NumPy.

The \code{ndarray} is the data structure at the heart of NumPy.
The data structure stores regularly strided homogeneous data types
in a contiguous block memory, allowing for the efficient representation
of $n$-dimensional data. \fixme{[How much detail do we want here?]}
More details about the data structure are given in ``The NumPy array:
a structure for efficient numerical computation.''\cite{vanderwalt2011numpy}.

The \emph{universal functions}, or more concisely, \emph{ufuncs},
are functions written in C that implement efficient looping over
NumPy arrays. An important feature of ufuncs is the built-in
implementation of \emph{broadcasting}.  For example, the function
\code{arctan2(x, y)} is a ufunc that accepts two values and computes
$\tan^{-1}(y/x)$.  When arrays are passed in as the arguments,
the ufunc will take care of looping over the dimensions of the inputs
in such a way that if, say, \code{x} is a 1-d array with length 3, and
\code{y} is a 2-d array with shape $2 \times 1$, the output will be
an array with shape $2 \times 3$.  The ufunc machinery takes care
of calling the function with all the appropriate combinations of
input array elements to complete the output array.
The elementary arithmetic operations of addition, multiplication, etc.,
are implemented as ufuncs, so broadcasting applies even in expressions
such as \code{x + y * z}.

\paragraph{\emph{Computing libraries.}}
NumPy provides a large library of functions for array manipulation
and scientific computing, including functions for: creating, reshaping,
concatenating, and padding arrays; searching, sorting and counting data
in arrays; computing elementary statistics, such as the mean, median,
variance, and standard deviation; file I/O; and more.

A suite of functions for computing the \emph{fast Fourier transform (FFT)}
and its inverse is provided.

NumPy's linear algebra library includes functions for: solving linear
systems of equations; computing various functions of a matrix, including
the determinant, the norm, the inverse, and the pseudo-inverse;
computing the Cholesky, eigenvalue, and singular value decompositions of a matrix;
and more.

The random number generator library in NumPy provides several alternative
\emph{bit stream generators} that provide the core function of generating
random integers.
A higher-level generator class that implements an assortment of
probability distributions is provided. It includes the beta, gamma
and Weibull distributions, the univariate and multivariate normal
distributions, and more.

A \emph{masked array} class, built on top of the \code{ndarray}, is
provided.  This class allows elements of an array to be masked,
indicating that the data is missing or unknown.

There is also a \code{matrix} class the implements 2-d arrays with
specialized behavior.  We mention it here for completeness, but the
use of \code{matrix} is discouraged.  The quirks of the \code{matrix}
class have proven to be troublesome, and with the addition of the
infix \code{@} operator to Python for matrix multiplication, there was no
longer a need for a separate \code{matrix} class, since one can
use a 2-d array instead.

\paragraph{\emph{Infrastructure libraries.}} NumPy provides utilities
for writing unit tests and for building Python packages.

The \code{testing} subpackage provides functions such as
\code{assert\_allclose(actual, desired)} that may be used in unit
test suites for code that uses NumPy arrays.

NumPy provides a subpackage includes functions and classes
that can be used by libraries that have NumPy as dependency,
to be used in the library's configuration code that installs the
library and that builds a package that can be used, for example,
on the PyPI web site.

\paragraph{\emph{F2PY.}}  The program \code{f2py} is a tool for
building NumPy-aware Python wrappers of Fortran functions.
NumPy itself does not use any Fortran code;  F2PY is part of NumPy
for historical reasons.
